Climbing

### 2021-07-27 manuscript/Rmd/draft.Rmd

## Other Scoring Methods

@@

### Why product of ranks instead of sum

@@

Using past competitions data, w are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).

Figure 3 is a scatterplot and correlation matrix between the ranks for the 2018 Youth Olympics Women's Qualification.

It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings. On the other hand, the relationship is not as strong for speed climbing. Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.

@@

```{r}
# wq %>% 
#   select(rank, speed:lead) %>% 
#   cor(method = "kendall") %>% 
#   data.frame()
```

@@

A single climber excluded changes things drastically, espcially order of medalists.

@@

>>>>

## Other Scoring Systems

@@

### Sum of ranks instead of product

@@

For our analysis on the relationship between the rankings of the events and the final result, we used data from the 2018 Youth Olympics Women's Qualification. Figure 3 is a scatterplot and correlation matrix between the ranks of the individual events and the final standings, with Kendall’s Tau (Kendall Rank Correlation Coefficient) as our measure of ordinal association between the quantities. It is evidently clear that there is a strong and positive correlation between the ranks of bouldering and lead climbing, and as a results, the standings of these two events are highly correlated with the final rankings. On the other hand, the correlation with the final rank is not as strong for speed climbing. Thus, speed climbers are facing a huge disadvantage in this scoring system, compared to those that are specialized in the other two concentrations.

This trend also holds for most of the past competitions.

@@

@@

A single climber excluded changes things drastically, especially order of medalists.

@@

Example from 2018 youth, women's final
dropping ranks 3 and 5 change the medalist order
